Hydra Battles in Coq

### 2020-11-21 doc/chapter-powers.tex
modifying the arguments of correctness predicates

\label{chain-correct-def}
\begin{Coqsrc}
Definition chain_correct_nat (c: chain) (n:nat) := 
  n <> 0 /\
  forall `(M:@EMonoid  A E_op E_one E_eq) (x:A), 
      chain_apply c x ==   x ^ n.

Definition chain_correct (c: chain) (p:positive) :=
  chain_correct_nat c (Pos.to_nat p). 
\end{Coqsrc}

@@

\begin{Coqsrc}
Ltac slow_chain_correct_tac :=
  match goal with 
      [ |- chain_correct ?c ?p ] =>
      let A := fresh "A" in
      let op := fresh "op" in
      let one := fresh "one" in
      let eqv := fresh "eqv" in
      let M := fresh "M" in
      let x := fresh "x"
      in  split;
        [discriminate | 
         unfold c, chain_apply, computation_eval; simpl;
         intros A op one eq M x; monoid_simpl M; reflexivity]
  end.

Example C7_ok : chain_correct C7 7.
Proof.
   slow_chain_correct_tac.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Example C87_ok : chain_correct C87 87.
Proof.
 Time  slow_chain_correct_tac. 
\end{Coqsrc}

@@

\begin{Coqsrc}
Ltac reflection_correct_tac :=
match goal with
[ |- chain_correct ?c ?n ] =>
 split; [try discriminate |
         let A := fresh "A"
         in let op := fresh "op"
         in let one := fresh "one" 
         in let E_eq := fresh "eq" 
         in let M := fresh "M"
         in let x := fresh "x" 
         in  (try unfold c); unfold chain_apply;
           simpl; red; intros  A op one E_eq M x;
           unfold computation_eval;simpl;
           monoid_eq_A A op one E_eq M;
           apply flatten_valid_2;try reflexivity
        ]
end. 
\end{Coqsrc}

@@

\begin{Coqsrc}
Example C87_ok : chain_correct C87 87.
Proof.
  Time reflection_correct_tac.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma param_correctness_nat :
 forall c:chain, parametric c ->  
               chain_correct_nat c  (the_exponent_nat c).
(* Proof omitted *)
\end{Coqsrc}

@@

Lemma param_correctness (c:chain) :
   parametric c -> 
   chain_correct c (the_exponent c). 
Proof.
  intros; rewrite  exponent_pos_of_nat; auto.
  red;  rewrite  exponent_pos2nat;auto.
  rewrite Pos2Nat.id,  <- exponent_pos2nat;auto.
  apply param_correctness_nat; auto.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Ltac param_chain_correct :=
match goal with 
[|- chain_correct ?c ?p ] => 
apply param_correctness; parametric_tac
end.

Lemma C87_ok' : chain_correct C87 87.
Time param_chain_correct.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition correct_generator (g : positive -> chain) :=
 forall p, chain_correct p (g p).
\end{Coqsrc}

@@

\begin{Coqsrc}
Remark F9_correct :chain_correct (F2C F9) 9.
Proof.
  apply param_correctness_pos;  lazy;  parametric_tac.
Qed.
\end{Coqsrc}

@@

Definition Fchain_correct  (f : Fchain) (n:nat) :=
  forall A `(M : @EMonoid A op E_one E_equiv) k (a:A),
    computation_execute op (f A k  a)==
    computation_execute op (k  (a ^ n)).
\end{Coqbad}

@@

\begin{Coqbad}
Theorem F3_correct : Fchain_correct F3 3.
Proof.  
  intros    A op E_one E_equiv M k  a ; simpl.
  monoid_simpl M.  
\end{Coqbad}

@@

\begin{Coqsrc}
Definition Fchain_correct_nat (f : Fchain) (n:nat) :=
 forall A `(M : @EMonoid A op E_one E_equiv) k
        (Hk :Fkont_proper M k)
        (a : A) ,
 computation_execute op (f A k  a) ==
 computation_execute op (k  (a ^ n)).

Definition Fchain_correct (f : Fchain) (p:positive) :=
 Fchain_correct_nat f (Pos.to_nat p).
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma F1_correct : Fchain_correct F1 1.
Proof.
  intros until M ; intros k Hk a ; unfold F1; simpl.
  apply Hk; monoid_simpl M; reflexivity.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Theorem F3_correct : Fchain_correct F3 3.
Proof. 
  intros until M; intros k Hk a; simpl.
\end{Coqsrc}

@@

\begin{Coqsrc}
Theorem F2_correct : Fchain_correct F2 2.
Proof. 
  intros until M; intros k Hk a; simpl;
  apply Hk;  monoid_simpl M;  reflexivity.
Qed.
\end{Coqsrc}

@@

Lemma Fcompose_correct_attempt :
  forall f1 f2 n1 n2, Fchain_correct f1 n1 ->
                      Fchain_correct f2 n2 ->
                      Fchain_correct (Fcompose f1 f2)
                                       (n1 * n2).

@@

\begin{Coqsrc}
Lemma  Fexp2_nat_correct (n:nat) : 
         Fchain_correct_nat (Fexp2_of_nat n) (2  ^ n).
Proof.
  induction n; cbn.
 - apply F1_correct.
 -  rewrite Nat.add_0_r;
   replace (2 ^ n + 2 ^ n)%nat with (2 * 2 ^n)%nat by  omega;
   apply Fcompose_correct_nat;auto.
   +  apply F2_correct.
   +  apply  Fexp2_nat_proper.
Qed.
\end{Coqsrc}

\begin{Coqsrc}
Lemma  Fexp2_correct (p:positive) : 
                           Fchain_correct (Fexp2 p) (2  ^ p).
(* Proof omitted *)

@@

Example F144:  {f : Fchain | Fchain_correct f 144 /\
                                Fchain_proper f}.
Proof.
 change 144 with ( (3 * 3) * (2 ^ 4))%positive.
 exists (Fcompose (Fcompose F3 F3) (Fexp2 4)); auto.
Defined.

@@

 \begin{Coqsrc}
Definition Kchain_correct_nat (kc : Kchain) (n p : nat) :=
  forall `(M : @EMonoid A op E_one E_equiv)
        (k : Kkont A), 
        Kkont_proper M k ->
        forall  (x : A) ,
              computation_execute op (kc  A k  x) ==
              computation_execute op (k  (x ^ n) (x ^ p)).

@@

\begin{Coqsrc}
Definition Kchain_correct (kc : Kchain) (n p : positive) :=
  Kchain_correct_nat kc (Pos.to_nat n) (Pos.to_nat p).

@@

Lemma k7_3_correct : Kchain_correct k7_3 7 3.
Proof.
  intros until M; intros; red; unfold k7_3; simpl.
  apply H; monoid_simpl M;  reflexivity.
Qed. 
\end{Coqsrc}

@@

Lemma K2F_correct :
  forall knp n p, Kchain_correct kc n p ->
                 Fchain_correct (K2F knp) n.
(* Proof omitted *)

@@

Lemma KFK_correct :
  forall (b q r : positive) (kbr : Kchain) (fq : Fchain),
    Kchain_correct kbr b r ->
    Fchain_correct fq q ->
    Kchain_proper kbr ->
    Fchain_proper fq -> 
    Kchain_correct (KFK kbr fq) (b * q + r) b.
(* Proof omitted *)

@@

Lemma KFF_correct :
forall (b q r : positive) (kbr : Kchain) (fq : Fchain),
Kchain_correct kbr b r ->
Fchain_correct fq q ->
Kchain_proper kbr ->
Fchain_proper fq -> Fchain_correct (KFF kbr fq) (b * q + r).
(* Proof omitted *)

@@

Lemma FK_correct : forall (p: positive) (Fp : Fchain),
                     Fchain_correct  Fp p ->
                     Fchain_proper Fp ->
                     Kchain_correct (FK Fp) p 1.
(* Proof omitted *)

@@

Lemma FFK_correct :
  forall (p q  : positive) (fp fq : Fchain),
    Fchain_correct fp p  ->
    Fchain_correct fq q ->
    Fchain_proper fp ->
    Fchain_proper fq -> Kchain_correct (FFK fp fq) (p * q ) p.
(* Proof omitted *)

@@

Lemma OK87 : Fchain_correct F87 87.
Proof.
 unfold F87; change 87 with (10 * (2 ^ 3) + 7)%positive.
 apply KFF_correct;auto.
 change 10 with (7 * 1 + 3); apply KFK_correct;auto.
 change 7 with (3 * 2 ^ 1 + 1)%positive;  apply KFK_correct;auto.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Theorem make_chain_correct : 
   forall p, chain_correct (make_chain p) p.  
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition correctness_statement (s: signature) : 
chain_type s -> Prop :=
match s  with
  | gen_F p => fun ch => Fchain_correct ch p
  | gen_K p d   => fun ch => Kchain_correct ch (p + d) p
end.

@@

Theorem make_chain_correct :
    forall p, chain_correct (make_chain p) p.
Proof.
 intro p; destruct (chain_gen_OK (gen_F p)).
 unfold make_chain; apply F2C_correct; apply H.
Qed.
\end{Coqsrc}

@@

%\subsection{Optimality and Relative Efficiency}

% \begin{project}[Proof techniques]
% \noindent

% \begin{enumerate}
% \item Improve automated proofs on types \texttt{positive} and \texttt{N}.
% \item Compare  \texttt{Program Fixpoint} and \texttt{Function} for
% writing \texttt{make\_chain}. Consider measure \emph{vs} well-founded 
% relations, mutual recursion, possibility of using sigma-types, etc.
% \item Chains are always associated with strictly positive exponents. 
% Thus, many lemmas about chain correctness  can be proved using semi-groups instead of
% monoids. Define type classes for semi-groups and use them whenever possible.
% \end{enumerate}  
% \end{project}

>>>>

\label{chain-correct-def}
\begin{Coqsrc}
Definition chain_correct_nat (n:nat) (c: chain) := 
  n <> 0 /\
  forall `(M:@EMonoid  A E_op E_one E_eq) (x:A), 
      chain_apply c x ==   x ^ n.

Definition chain_correct (p:positive) (c: chain) :=
  chain_correct_nat c (Pos.to_nat p). 
\end{Coqsrc}

@@

\begin{Coqsrc}
Ltac slow_chain_correct_tac :=
  match goal with 
      [ |- chain_correct ?p ?c ] =>
      let A := fresh "A" in
      let op := fresh "op" in
      let one := fresh "one" in
      let eqv := fresh "eqv" in
      let M := fresh "M" in
      let x := fresh "x"
      in  split;
        [discriminate | 
         unfold c, chain_apply, computation_eval; simpl;
         intros A op one eq M x; monoid_simpl M; reflexivity]
  end.

Example C7_ok : chain_correct 7 C7.
Proof.
   slow_chain_correct_tac.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Example C87_ok : chain_correct 87 C87.
Proof.
 Time  slow_chain_correct_tac. 
\end{Coqsrc}

@@

\begin{Coqsrc}
Ltac reflection_correct_tac :=
match goal with
[ |- chain_correct ?n ?c ] =>
 split; [try discriminate |
         let A := fresh "A"
         in let op := fresh "op"
         in let one := fresh "one" 
         in let E_eq := fresh "eq" 
         in let M := fresh "M"
         in let x := fresh "x" 
         in  (try unfold c); unfold chain_apply;
           simpl; red; intros  A op one E_eq M x;
           unfold computation_eval;simpl;
           monoid_eq_A A op one E_eq M;
           apply flatten_valid_2;try reflexivity
        ]
end. 
\end{Coqsrc}

@@

\begin{Coqsrc}
Example C87_ok : chain_correct 87 C87.
Proof.
  Time reflection_correct_tac.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma param_correctness_nat :
 forall c:chain, parametric c ->  
               chain_correct_nat (the_exponent_nat c)  c.
(* Proof omitted *)
\end{Coqsrc}

@@

Lemma param_correctness (c:chain) :
   parametric c -> 
   chain_correct (the_exponent c) c. 
Proof.
  intros; rewrite  exponent_pos_of_nat; auto.
  red;  rewrite  exponent_pos2nat;auto.
  rewrite Pos2Nat.id,  <- exponent_pos2nat;auto.
  apply param_correctness_nat; auto.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Ltac param_chain_correct :=
match goal with 
[|- chain_correct ?p ?c ] => 
apply param_correctness; parametric_tac
end.

Lemma C87_ok' : chain_correct 87 C87.
Time param_chain_correct.
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition correct_generator (g : positive -> chain) :=
 forall p, chain_correct (g p) p.
\end{Coqsrc}

@@

\begin{Coqsrc}
Remark F9_correct :chain_correct 9 (F2C F9).
Proof.
  apply param_correctness_pos;  lazy;  parametric_tac.
Qed.
\end{Coqsrc}

@@

Definition Fchain_correct  (n:nat) (f : Fchain) :=
  forall A `(M : @EMonoid A op E_one E_equiv) k (a:A),
    computation_execute op (f A k  a)==
    computation_execute op (k  (a ^ n)).
\end{Coqbad}

@@

\begin{Coqbad}
Theorem F3_correct : Fchain_correct 3 F3.
Proof.  
  intros    A op E_one E_equiv M k  a ; simpl.
  monoid_simpl M.  
\end{Coqbad}

@@

\begin{Coqsrc}
Definition Fchain_correct_nat (n:nat) (f : Fchain) :=
 forall A `(M : @EMonoid A op E_one E_equiv) k
        (Hk :Fkont_proper M k)
        (a : A) ,
 computation_execute op (f A k  a) ==
 computation_execute op (k  (a ^ n)).

Definition Fchain_correct (p:positive) (f : Fchain) :=
 Fchain_correct_nat (Pos.to_nat p) f.
\end{Coqsrc}

@@

\begin{Coqsrc}
Lemma F1_correct : Fchain_correct 1 F1.
Proof.
  intros until M ; intros k Hk a ; unfold F1; simpl.
  apply Hk; monoid_simpl M; reflexivity.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Theorem F3_correct : Fchain_correct 3 F3.
Proof. 
  intros until M; intros k Hk a; simpl.
\end{Coqsrc}

@@

\begin{Coqsrc}
Theorem F2_correct : Fchain_correct 2 F2.
Proof. 
  intros until M; intros k Hk a; simpl;
  apply Hk;  monoid_simpl M;  reflexivity.
Qed.
\end{Coqsrc}

@@

Lemma Fcompose_correct_attempt :
  forall f1 f2 n1 n2, Fchain_correct n1 f1 ->
                      Fchain_correct n2 f2 ->
                      Fchain_correct (n1 * n2) (Fcompose f1 f2).

@@

\begin{Coqsrc}
Lemma  Fexp2_nat_correct (n:nat) : 
         Fchain_correct_nat (2  ^ n) (Fexp2_of_nat n).
Proof.
  induction n; cbn.
 - apply F1_correct.
 -  rewrite Nat.add_0_r;
   replace (2 ^ n + 2 ^ n)%nat with (2 * 2 ^n)%nat by  omega;
   apply Fcompose_correct_nat;auto.
   +  apply F2_correct.
   +  apply  Fexp2_nat_proper.
Qed.
\end{Coqsrc}

\begin{Coqsrc}
Lemma  Fexp2_correct (p:positive) : 
                           Fchain_correct (2  ^ p) (Fexp2 p).
(* Proof omitted *)

@@

Example F144:  {f : Fchain | Fchain_correct 144 f /\
                                Fchain_proper f}.
Proof.
 change 144 with ( (3 * 3) * (2 ^ 4))%positive.
 exists (Fcompose (Fcompose F3 F3) (Fexp2 4)); auto.
Defined.

@@

 \begin{Coqsrc}
Definition Kchain_correct_nat (n p : nat) (kc : Kchain) :=
  forall `(M : @EMonoid A op E_one E_equiv)
        (k : Kkont A), 
        Kkont_proper M k ->
        forall  (x : A) ,
              computation_execute op (kc  A k  x) ==
              computation_execute op (k  (x ^ n) (x ^ p)).

@@

\begin{Coqsrc}
Definition Kchain_correct (n p : positive) (kc : Kchain) :=
  Kchain_correct_nat  (Pos.to_nat n) (Pos.to_nat p) kc.

@@

Lemma k7_3_correct : Kchain_correct 7 3 k7_3.
Proof.
  intros until M; intros; red; unfold k7_3; simpl.
  apply H; monoid_simpl M;  reflexivity.
Qed. 
\end{Coqsrc}

@@

Lemma K2F_correct :
  forall knp n p, Kchain_correct kc n p ->
                 Fchain_correct (K2F n) knp.
(* Proof omitted *)

@@

Lemma KFK_correct :
  forall (b q r : positive) (kbr : Kchain) (fq : Fchain),
    Kchain_correct b r  kbr ->
    Fchain_correct q fq ->
    Kchain_proper kbr ->
    Fchain_proper fq -> 
    Kchain_correct  (b * q + r) b (KFK kbr fq).
(* Proof omitted *)

@@

Lemma KFF_correct :
forall (b q r : positive) (kbr : Kchain) (fq : Fchain),
Kchain_correct  b r kbr ->
Fchain_correct q fq ->
Kchain_proper kbr ->
Fchain_proper fq -> Fchain_correct (b * q + r) (KFF kbr fq).
(* Proof omitted *)

@@

Lemma FK_correct : forall (p: positive) (Fp : Fchain),
                     Fchain_correct  p Fp ->
                     Fchain_proper Fp ->
                     Kchain_correct p 1 (FK Fp) .
(* Proof omitted *)

@@

Lemma FFK_correct :
  forall (p q  : positive) (fp fq : Fchain),
    Fchain_correct p fp  ->
    Fchain_correct q fq ->
    Fchain_proper fp ->
    Fchain_proper fq -> Kchain_correct  (p * q ) p (FFK fp fq).
(* Proof omitted *)

@@

Lemma OK87 : Fchain_correct 87 F87.
Proof.
 unfold F87; change 87 with (10 * (2 ^ 3) + 7)%positive.
 apply KFF_correct;auto.
 change 10 with (7 * 1 + 3); apply KFK_correct;auto.
 change 7 with (3 * 2 ^ 1 + 1)%positive;  apply KFK_correct;auto.
Qed.
\end{Coqsrc}

@@

\begin{Coqsrc}
Theorem make_chain_correct : 
   forall p, chain_correct p (make_chain p).  
\end{Coqsrc}

@@

\begin{Coqsrc}
Definition correctness_statement (s: signature) : 
chain_type s -> Prop :=
match s  with
  | gen_F p => fun ch => Fchain_correct p ch
  | gen_K p d   => fun ch => Kchain_correct (p + d) p ch
end.

@@

Theorem make_chain_correct :
    forall p, chain_correct p (make_chain p).
Proof.
 intro p; destruct (chain_gen_OK (gen_F p)).
 unfold make_chain; apply F2C_correct; apply H.
Qed.
\end{Coqsrc}

@@

\begin{project}[Some improvements]
 \begin{enumerate}
\item Improve automated proofs on types \texttt{positive} and \texttt{N}.
\item Compare  \texttt{Program Fixpoint} and \texttt{Function} for
writing \texttt{make\_chain}. Consider measure \emph{vs} well-founded 
relations, mutual recursion, possibility of using sigma-types, etc.
\item Chains are always associated with strictly positive exponents. 
Thus, many lemmas about chain correctness  can be proved using semi-groups instead of
monoids. Define type classes for semi-groups and use them whenever possible.
\end{enumerate}  
\end{project}

% \end{project}
